---
title: "QMSS Final Project"
output: html_notebook
---
The Physician Compare website was created by the Centers for Medicare & Medicaid Services (CMS) in December 2010 as required by the Affordable Care Act (ACA) of 2010 to help patients assess and find doctors and hospitals. This dataset contains the information supplied to patients via that website, including patient satisfaction surveys and performance scores across over 100 metrics.

Experience:  
Each unique measure title has a different method of determining the measure score, but it's scaled by percent, where higher is better performance in that category/title.


```{r, include = F}
invisible(library(dplyr))
```

```{r}
experience <- read.csv("Physician_Compare_2015_Group_Public_Reporting___Patient_Experience.csv", 
                       header = T, na.strings=c("", "NA"))
group_score <- read.csv("Physician_Compare_2015_Group_Public_Reporting___Performance_Scores.csv",
                        header = T, na.strings=c("", "NA"))
indiv_score <- read.csv("Physician_Compare_2015_Individual_EP_Public_Reporting___Performance_Scores.csv", 
                        header = T, na.strings=c("", "NA"))
national <- read.csv("Physician_Compare_National.csv", 
                     header = T, na.strings=c("", "NA"))
```

```{r}
head(experience)
```

```{r}
# will want to remove inverse scores 
# check if Participating.in.PQRS affects measure identifier and the proportion of them

# reporting mechanism = collection_type
# check footnotes?
head(group_score)
```


```{r}
head(indiv_score)
```

```{r}

```


```{r}
head(national)
```


```{r}
colSums(!(is.na(national["All.secondary.specialties"])))
dim(national)
```

