---
title: "QMSS Final Project"
output:
  pdf_document: default
  html_notebook: default
---

The Physician Compare website was created by the Centers for Medicare & Medicaid Services (CMS) in December 2010 as required by the Affordable Care Act (ACA) of 2010 to help patients assess and find doctors and hospitals. This dataset contains the information supplied to patients via that website, including patient satisfaction surveys and performance scores across over 100 metrics.

Looking at individual physician scores:  

* MIPS
* Performance by measure category
* Organization MIPS

Possible problems to solve:  
 
* Prediction of whether to see a physician based on threshold
* Predict physician score based on all factors (predict MIPS?)
    + classification of whether you should go to a doctor based on MIPS score
    + train on several classification methods to analyze best approach for this data


```{r, include = F}
invisible(library(dplyr))
invisible(library(stringr))
invisible(library(tidyr))
invisible(library(dmm))
```


```{r}
dr_scores <- read.csv("full_doctor_scoring.csv", sep = ",", na = "NA")
head(dr_scores)
```

Potential Methods for Binary Classification:  
Using overall MIPS for individuals where MIPS>= 75, the positive payment adjustment threshold.

* could apply spline to other MIPS, ACI scorings since they're somewhat discrete in nature.  
    + ACI >= 0 : clinician reported ACI category
    + ACI >= 50: clinician achieved base score for ACI
    + MIPS < 30: Negative Payment Adjustment
* Predictive MIPS >= 75, essentially.
* Methods to try:  
    + glmnet for binary classification (elastic model/penalized logit)
    + glm logit model with polynomials?
    + tree model if we can make it work? (Single Tree, Random Forest, Boosting, Dbarts???)
    + PLSDA or LDA
* PCA to look at similar variables? like an inverted version
    





